---
title: "Cleaning and wrangling data"
format: html
---

## Set up
```{r}
#| warning: FALSE
#| message: FALSE
library(dplyr)
library(tidyr)
library(readr)
```

## Read data

```{r}
#| message: FALSE
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

```{r}
#| echo: false
#| eval: false
colnames(catch_original)
summary(catch_original)
head(catch_original)
unique(catch_original$Region)
```

## Using `select()` to remove or select column

```{r}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
```

Quality check using `glimpse()` (similar to `str()`)

```{r}
glimpse(catch_data)
```

Shows that the Chinook column read in as character string when it should be numeric.
Override Chinook column using `mutate()` to force it to be numeric.

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))
```

Get a warning that an NA was introduced by forcing column to be numeric (something wasn't numeric).
Find out where the NA was introduced.

```{r}
i <- which(is.na(catch_clean$Chinook))
i
catch_data[i,]
```

There was a letter "I" instead of a number in one cell. 
Replace it using `mutate()` and force column to be numeric.

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(condition = Chinook == "I",
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))

glimpse(catch_clean)
```

## Reshaping DF with `pivot_longer()` and `pivot_wider()`

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols=-c(Region, Year),
               names_to = "species",
               values_to = "catch")
head(catch_long)
```

## Renaming a column using `rename()`

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousand = catch)
colnames(catch_long)
```

## Adding a column using `mutate()` and deleting another column using `select()`

```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousand*1000) %>% 
  select (-catch_thousand)

head(catch_long)
```

## Using a 'Split-Apply-Combine' approach to get mean catch by region

```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(mean_catch=mean(catch))

head(mean_region)

n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(n=n())

head(n_region)
```

## Using `filter()`

```{r}
sse_catch <- catch_long %>% 
  filter(Region == "SSE")

head(sse_catch)
unique(sse_catch)

sse_nse_catch <- catch_long %>% 
  filter(Region %in% c("SSE", "NSE"))

unique(sse_nse_catch)
```

