---
title: "Mini Literate Analysis"
author: "Ingrid Slette"
format: html
---

## Introduction
### About the data
This data set contains North Pole Environmental Observatory Bottle Chemistry data.

The data can be downloaded from the [Arctic Data Center repository](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X).

### Setup
```{r}
library(readr)
library(here)
```

### Read the data

```{r}
#| warning: false
bg_chem <- read_csv(here("data/BGchem2008data.csv"))
```

Explore the data
```{r}
#| eval: false
colnames(bg_chem)
str(bg_chem)
summary(bg_chem)
unique(bg_chem$Station)
```

## Analysis
### Calculate summary statistics
```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

### Calculate mean Redfield ratio
```{r}
ratio <- (nitrate + nitrite + amm)/phos
```

### Plot the Redfield ratio
```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```


## Result

The Redfield ratio of this data is `r round(ratio, digits =2)`.
