---
title: "R Practice: Tidy Data and Joins"
format: html
---

## Set up

```{r}
#| warning: FALSE
#| message: FALSE
library(readr)
library(dplyr)
library(lubridate) # for bonus question
```

## Read data

```{r}
#| message: FALSE
bird_observations <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=53edaa7a0e083013d9bf20322db1780e")

sites <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=81bf72420e69077097fb0790dcdc63a6")

surveys <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=b2466fa5cb5ed7ee1ea91398fc291c59")

taxalist <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=58f863b7e3066e68536a9cacdc7bd58e")
```

## Explore data

```{r}
#| echo: false
glimpse(bird_observations)
str(bird_observations)
unique(bird_observations)
head(bird_observations)
summary(bird_observations)
glimpse(surveys)
head(surveys)
glimpse(sites)
head(sites)
glimpse(taxalist)
head(taxalist)
```
## Subset bird observations table (only species BHCO and RWBL and sites LI-W and NU-C)

```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))
```

## Left join bird observations subset with sites df

```{r}
#| echo: false
colnames(birds_subset)
colnames(sites)
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id")
```

## Full join bird observations subset and sites df

```{r}
birds_full <- birds_subset %>% 
  full_join(y = sites, by = "site_id")
```

## Inner join bird observations subset and taxalist df

```{r}
birds_inner <- birds_subset %>% 
  inner_join(y = taxalist, by = "species_id")
```

## Left join bird observations subset and taxalist

```{r}
birds_inner_left <- birds_subset %>% 
  left_join(y = taxalist, by = "species_id")
```

## Rename 'notes' column in bird observations (so doens't conflict with 'notes' column in surveys df)

```{r}
bird_observations <- bird_observations %>% 
  rename(bird_obs_notes = notes)
```

## Join sites df, surveys df, and taxalist df to bird observations df and keep only the  survey_date, common_name, park_name, bird_count, and observer columns

```{r}
bird_obs_subset <- bird_observations %>% 
  full_join(y = taxalist, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = c("site_id", "survey_id")) %>%  
  select(survey_date, common_name, park_name, bird_count, observer)
```

## Check values of observer

```{r}
unique(bird_obs_subset$observer)
```

## Change so that all observer names are formatted the same way 

```{r}
bird_obs_subset <- bird_obs_subset %>% 
  mutate(observer = if_else(condition = observer == "Josh Burns", 
                            true = "J. Burns", 
                            false = observer))

unique(bird_obs_subset$observer)
```

